---
title: "Coursera Getting and Cleaning Data Final Project"
author: "Caio Costa"
date: "10/18/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Let's start by reading the tables

```{r reading}
#Reading stuff, fixed width file, 16 characters, 561 features
x_train <- as_tibble(read.fwf("train/X_train.txt", widths = rep(c(16),561)))
x_test <- as_tibble(read.fwf("test/X_test.txt", widths = rep(c(16),561)))
# Single column file, no headers
y_train <- as_tibble(read.csv("train/Y_train.txt", header = FALSE))
y_test <- as_tibble(read.csv("test/Y_test.txt",header = FALSE))
#Reading the subject, will need those in the final step. Believe me
subject_train <- as_tibble(read.csv("train/subject_train.txt",header = FALSE))
subject_test <- as_tibble(read.csv("test/subject_test.txt",header = FALSE))
#Reading the feature names
labels <- read.csv("features.txt",header = FALSE, sep = " ")
# Labeling the features
names(x_train) <- labels$V2
names(x_test) <- labels$V2
names(y_train) <- c("activity_code")
names(y_test) <- c("activity_code")
names(subject_train) <- c("subject")
names(subject_test) <- c("subject")
#Gluing stuff together
train <-bind_cols(x_train,y_train,subject_train)
test <- bind_cols(x_test,y_test,subject_test)
#Just leaving a marker in case I need in the feature
train <- mutate(train, type = "train")
test <- mutate(test, type = "test")
#Gluing some more
data <- bind_rows(train,test)
#Voila
print(data)
```

Now let's extract those mean fields (pun intended)

```{R slicing}
#Slicing just means and std (and activity code and type and subject)
data2 <- data %>%
  select(ends_with("mean()"), ends_with("std()"), `activity_code`, type, subject)
print(data2)
```

Now let's read that activity_labels file and figure out what dat code means

```{R activities}
#Reading activities descriptions and joining to data2
activities <- read.csv("activity_labels.txt",header = FALSE, sep = " ")
names(activities) <- c("activity_code","activity_description")
data3 <- inner_join(data2,activities,by = c("activity_code"))
print(select(data3,activity_code,activity_description))
```

I am pretty comfortable with my variable names right now. Does anyone get to this point without naming them? 

Last step. Second data set? I am already in the count of 4. But okay. Let's average those by activity and subject.

```{R grouping}
data4 <- data3 %>%
  select(-type, -activity_code) %>%
  group_by(subject,activity_description) %>%
  summarize_all(funs(mean))
print(data4)
```
