---
title: "Course project 1"
author: "Caio Costa"
date: "10/27/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading and fixing data

Let's take a glimpse of the data, making sure it is fine

```{r gilmpse}
data <- read.csv("household_power_consumption.txt", sep = ';', stringsAsFactors = FALSE) #hoped this would automatically finds the right type, but it defaulted to <chr>
head(data)
```

Everything looks good from here. Except it decided all values are <chr>. Let's fix that

```{r converting}
data$Datetime <- as.POSIXct(strptime(paste(data$Date,data$Time), format = '%d/%m/%Y %H:%M:%S')) #Need to convert to POSIXct, because strptime output is POSIXlt, which is a list and don't do well on a dataframe

data$Global_active_power <- as.numeric(data$Global_active_power)
data$Global_reactive_power <- as.numeric(data$Global_reactive_power)
data$Voltage <- as.numeric(data$Voltage)
data$Global_intensity <- as.numeric(data$Global_intensity)
data$Sub_metering_1 <- as.numeric(data$Sub_metering_1)
data$Sub_metering_2 <- as.numeric(data$Sub_metering_2)
data$Sub_metering_3 <- as.numeric(data$Sub_metering_3)
head(data)
```

Way better!

Now let's just narrow down to the dates in question

```{r filtering}
data <- subset(data, Datetime  >= '2007-02-01' & Datetime < '2007-02-03')
#Just to make sure
c(min(data$Datetime), max(data$Datetime))
```

Awesome, seems like we filtered just the data we need!

# First plot

First plot looks like a histogram. It seems to show how many times the Global Active Power was in a certain range.

```{R plot1}
hist(data$Global_active_power, 
     col = "red", 
     main = "Global Active Power", 
     xlab = "Global Active Power (kilowatts)")
```

This is awesome, that plot looks like what the exercise asked. Let's save it on a png

```{R save_plot1}
png(filename = 'plot1.png', width = 480, height = 480, bg = NA) #bg = NA gives us the nice png proeperty of transparent background
hist(data$Global_active_power, 
     col = "red", 
     main = "Global Active Power", 
     xlab = "Global Active Power (kilowatts)")
dev.off()
```

# Second plot

Hmm, this seems to be just a minute by minute line of Global Active Power. Let's do this!

```{R plot2}
plot(x = data$Datetime, 
     y = data$Global_active_power, 
     type = 'l', #we want lines, not dots
     ylab = "Global Active Power (kilowatts)",
     xlab = "")
```

YES! I knew all that work of setting the column types would yield good results! Look how easy and beautiful plotting that was!

Now saving to a png file

```{R save_plot2}
png(filename = 'plot2.png', width = 480, height = 480, bg = NA)
plot(x = data$Datetime, 
     y = data$Global_active_power, 
     type = 'l', 
     ylab = "Global Active Power (kilowatts)",
     xlab = "")
dev.off()
```

# Third plot

Getting fancy, we have legends now that tells us what we are plotting. Which also means we have to plot 3 things at the same time! Okay we can do this

```{R plot3}
plot(x = data$Datetime, 
     y = data$Sub_metering_1, 
     type = "l",
     col = "black",
     xlab = "",
     ylab = "Energy sub metering")
lines(x = data$Datetime, 
     y = data$Sub_metering_2, 
     col = "red")
lines(x = data$Datetime, 
     y = data$Sub_metering_3, 
     col = "blue")
legend("topright", 
       legend = c("Sub_metering_1","Sub_metering_2","Sub_metering_3"),
       lwd = c(2,2), #because we want lines on the legend
       col =c("black","red","blue"), #and we want them in the right color
       bty = "n") #bty = "n" gets rid of the legend box borders
```

Woohooo it works!

Just need to save it.

```{R save_plot3}
png(filename = 'plot3.png', width = 480, height = 480, bg = NA)
plot(x = data$Datetime, 
     y = data$Sub_metering_1, 
     type = "l",
     col = "black",
     xlab = "",
     ylab = "Energy sub metering")
lines(x = data$Datetime, 
     y = data$Sub_metering_2, 
     col = "red")
lines(x = data$Datetime, 
     y = data$Sub_metering_3, 
     col = "blue")
legend("topright", 
       legend = c("Sub_metering_1","Sub_metering_2","Sub_metering_3"),
       lwd = c(2,2),
       col =c("black","red","blue"),
       bty = "n")
dev.off()
```

# Fourth plot

It is the final challenge. 4 plots. 2 of them we already plotted, so that should be easy. Let's get the other 2 right. Voltage over time and Global_reactive_power over time.

```{r plot4}
par(mfrow = c(2,2)) #Set device to expect 2x2 matrix of plots
#Top left
plot(x = data$Datetime, 
     y = data$Global_active_power, 
     type = 'l', 
     ylab = "Global Active Power",
     xlab = "")
#Top right
plot(x = data$Datetime, 
     y = data$Voltage, 
     type = 'l', 
     ylab = "Voltage",
     xlab = "datetime")
#Botton left
plot(x = data$Datetime, 
     y = data$Sub_metering_1, 
     type = "l",
     col = "black",
     xlab = "",
     ylab = "Energy sub metering")
lines(x = data$Datetime, 
     y = data$Sub_metering_2, 
     col = "red")
lines(x = data$Datetime, 
     y = data$Sub_metering_3, 
     col = "blue")
legend("topright", 
       legend = c("Sub_metering_1","Sub_metering_2","Sub_metering_3"),
       lwd = c(2,2),
       col =c("black","red","blue"),
       bty = "n")
#Bottom right
plot(x = data$Datetime, 
     y = data$Global_reactive_power, 
     type = 'l', 
     ylab = "Global_reactive_power",
     xlab = "datetime")
```

Piece of cake. Now saving it

```{R save_plot4}
png(filename = 'plot4.png', width = 480, height = 480, bg = NA)
par(mfrow = c(2,2))
#Top left
plot(x = data$Datetime, 
     y = data$Global_active_power, 
     type = 'l', 
     ylab = "Global Active Power",
     xlab = "")
#Top right
plot(x = data$Datetime, 
     y = data$Voltage, 
     type = 'l', 
     ylab = "Voltage",
     xlab = "datetime")
#Botton left
plot(x = data$Datetime, 
     y = data$Sub_metering_1, 
     type = "l",
     col = "black",
     xlab = "",
     ylab = "Energy sub metering")
lines(x = data$Datetime, 
     y = data$Sub_metering_2, 
     col = "red")
lines(x = data$Datetime, 
     y = data$Sub_metering_3, 
     col = "blue")
legend("topright", 
       legend = c("Sub_metering_1","Sub_metering_2","Sub_metering_3"),
       lwd = c(2,2),
       col =c("black","red","blue"),
       bty = "n")
#Bottom right
plot(x = data$Datetime, 
     y = data$Global_reactive_power, 
     type = 'l', 
     ylab = "Global_reactive_power",
     xlab = "datetime")
dev.off()
```

That is all folks!